On The Impact of Atoms of Confusion in JavaScript Code

### 2021-10-21 2_Background.tex
ajustes no texto necessarios para o limite de 10 paginas

Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms in \clang. They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend. Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms. The authors argue that this method is important to complement a quantitative analysis. Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''}. We follow a similar approach as we augment our quantitative analysis with interviews.

>>>>

Oliveira et al.~\cite{TheEyesDoNotLie} conducted an experiment using an eye-tracker with 30 participants involving three atoms in \clang. They found that code with atoms of confusion requires more time from participants to predict the output and more visual effort to comprehend. Gopstein et al., in another paper~\cite{ThinkingAloud} further explored the original atoms by additionally conducting interviews and having programmers discuss among themselves the studied atoms. The authors argue in favor of complementing a quantitative analysis. Indeed, their study revealed findings such as a \textit{``correct evaluation of an atom might not mean that a programmer understood its meaning''}. Here, we also complement the quantitative analysis with interviews.

### 2021-10-21 3_Method.tex
ajustes no texto necessarios para o limite de 10 paginas

We decided not to provide feedback about the time students took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey in a social media platform, if we gave respondents instant feedback, they might post comments on particular atoms, therefore interfering with future participants' thought process.

@@

\begin{table}[htb!]
  \centering
  \caption{Demographic information of the participants}
\begin{scriptsize}  
\begin{tabular}{clcl}
\toprule
ID & Education & JS Experience & Other Languages \\ \midrule 
P1 & BSc Degree & 9 years & Java, PHP, C, Go  
\\ 
P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
\\ 
P3 & BSc Degree & 4 years & Java, C
\\ 
P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
\\ 
P5 & Master Student & 3 years & Python, SQL
\\ 
P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\# and R
\\ 
P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
\\ 
P8 & PhD Degree & 2 years & Java, Python
\\ 
P9 & BSc Degree  & 5 years & PHP
\\ 
P10 & Master Student & 4 years & Java, C\# and Python
\\ 
P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
\\ 
P12 & Master Degree & 1 year & C
\\ 
P13 & Master Degree & 13 years & Java, PHP
\\ 
P14 & Master Student & 3 years & C, Python, Ruby
\\ 
P15 &  BSc degree  & 2 years & Java, PHP
\\ \bottomrule
\end{tabular}
\end{scriptsize}
    \label{pinterview}
\end{table}

>>>>

We decided not to provide feedback about the time students took to answer each question. We also did not tell them whether their answers were correct or not. Our main concern was to avoid introducing bias for future respondents. Since we posted the survey in a social media platform, possible threats could have arisen if we gave respondents instant feedback.

@@

\begin{table}[htb!]
  \centering
  \caption{Demographic information of the participants}
\begin{scriptsize}  
\begin{tabular}{clcl}
\toprule
ID & Education & JS Experience & Other Languages \\ \midrule 
P1 & BSc Degree & 9 years & Java, PHP, C, Go  
\\ 
P2 & HS Degree & 3 years & Python, Go, Dart, Lua, C++, C\#
\\ 
P3 & BSc Degree & 4 years & Java, C
\\ 
P4 & Undergraduate & 3 years & Python, C, C++, Java, Go
\\ 
P5 & Master Student & 3 years & Python, SQL
\\ 
P6 & Bsc Degree & 15  years  & Java, PHP, C, Python , Ruby, C\#
\\ 
P7 & BSc Degree  & 6 years & C, C++, Java, Assembly, Kotlin
\\ 
P8 & PhD Degree & 2 years & Java, Python
\\ 
P9 & BSc Degree  & 5 years & PHP
\\ 
P10 & Master Student & 4 years & Java, C\# and Python
\\ 
P11 & Master Student & 4 years & Java, Erlang, C\#, Cobol
\\ 
P12 & Master Degree & 1 year & C
\\ 
P13 & Master Degree & 13 years & Java, PHP
\\ 
P14 & Master Student & 3 years & C, Python, Ruby
\\ 
P15 &  BSc degree  & 2 years & Java, PHP
\\ \bottomrule
\end{tabular}
\end{scriptsize}
    \label{pinterview}
\end{table}

### 2021-10-21 4_Result.tex
ajustes no texto necessarios para o limite de 10 paginas

{\color{blue}Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness. The findings suggest that \emph{participant experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}). Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform. For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets. It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code.} {\color{red}We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.} Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.

@@

{\color{red}Comparing to previous research, three 
atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
for \clang and \cpplang do not lead to a statistically significant impact
on correctness in our study: Logic as Control Flow,
Conditional Operator, and Pre-Increment. Our results
support the Gopstein et al. findings for the remaining
atoms. Considering the work of Langhout and Aniche~\cite{Langhout:2021:ACJ},
their work investigates six atom candidates for Java that
we also explore here for JavaScript. Among these atom
candidates, we achieve similar conclusions with their
study for four atom candidates: Post Increment,
Conditional Operator, Arithmetic as Logic, and
Omitted Curly Braces---both studies confirmed only the former
as an atom of confusion. Our findings diverged from
the work of Langhout and Aniche for two atoms: Pre Increment and
Logic as Control Flow.}

@@

One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving deep prototype chains. This is a core feature of the language, and most high-level tools and frameworks abstract it away.
%Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
% When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern---potentially leading to several levels of nested function calls---as extremely difficult to assimilate. One of the respondents stated:

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript. Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern. We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).  As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code. The following section presents the results of our mining software repository effort, whose goal is to reveal how frequent atom candidates are in open source code.

>>>>

Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either \emph{participant education} or \emph{participant experience} impact correctness. The findings suggest that \emph{participant experience} impacts on the results related to correctness (see Figure~\ref{fig:correctness-over-experience}). Even though the presence of atoms of confusion reduces the number of correct answers to all \emph{experience groups}, this effect is not uniform. For instance, novice developers provide a higher number of wrong answers for the confusing version of the code (57.8\%), while developers with more than four years of experience provide more than 70\% of correct answers even when evaluating the confusing version of the code snippets. It is also important to note that, independently of the \emph{experience group}, developers tend to provide more than 84\% of correct answers while predicting the outputs for the clean versions of the code. We replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience. Differently from \emph{participant experience}, the factor \emph{participant education} does not change substantially the correctness of the answers.

@@

Comparing to previous research, three 
atoms of confusion that Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} confirmed
for \clang and \cpplang do not lead to a statistically significant impact
on correctness in our study: Logic as Control Flow,
Conditional Operator, and Pre-Increment. Our results
support the Gopstein et al. findings for the remaining
atoms. Considering the work of Langhout and Aniche~\cite{Langhout:2021:ACJ},
their work investigates six atom candidates for Java that
we also explore here for JavaScript. Among these atom
candidates, we achieve similar conclusions with their
study for four atom candidates: Post Increment,
Conditional Operator, Arithmetic as Logic, and
Omitted Curly Braces---both studies confirmed only the former
as an atom of confusion. Our findings diverged from
the work of Langhout and Aniche for two atoms: Pre Increment and
Logic as Control Flow.

@@

One of the participants mentioned that the use of JavaScript's prototype-based inheritance can make it difficult to understand code, particularly when involving deep prototype chains.
%This is a core feature of the language, and most high-level tools and frameworks abstract it away.
%Although this was only mentioned by one practitioner, this is an important remark, as true understanding of JavaScript software necessarily involves understanding the concept of prototypes.
% When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern, which can lead to several levels of nested function calls, as extremely difficult to assimilate.
When asked about particular JavaScript constructs or patterns that can make code difficult to understand, three participants cited the callback pattern---potentially leading to several levels of nested function calls---as extremely difficult to assimilate. One of the respondents stated:

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript.
%Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.
We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).  As future work, we aim at investigating whether or not these atom candidates are more likely to introduce confusion in JavaScript code. The following section presents the results of our mining software repository effort, whose goal is to reveal how frequent atom candidates are in open source code.

### 2021-10-21 main.tex
rephrase one of the interview questions
ajustes no texto necessarios para o limite de 10 paginas

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: considering 10 atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects and two of them occur more than once for every 100 lines of code in the dataset. 
  %: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively.
  Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

% Our investigation also reveals additional JavaScript constructs that might hinder even experienced developers from understanding JavaScript code, such as the use of callbacks functions. Altogether...

%Evolving legacy code is a challenging task, particularly when the code has been poorly written---which imposes a significant cognitive load on developers when trying to familiarize themselves with a legacy codebase. Previous research has investigated the impact of small code patterns that contribute to misunderstanding the source code written in C and C++. However, little is known about the influence of these small code patterns (hereafter atoms of confusion) on understanding dynamic languages (such as JavaScript). In this work, we present the results of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. Our goal is to investigate the impact of atoms of confusion on understanding code written in JavaScript --- a dynamically typed language whose popularity is growing in the most different application domains (from web programming to backend and IOT development). From our survey and interviews, we confirm that atoms of confusion lead to code that is hard to understand: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in 6 out of 9 atom candidates. Also, the results of our MSR effort reveal that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects---including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Our investigation also reveals additional JavaScript constructs and idioms that might hinder developers from understanding JavaScript code. For instance, even experienced JavaScript developers agree that it is hard to understand code that uses callback functions, even though we found widespread use of callbacks in our projects dataset. Altogether, our findings might help developers better understand the implications of atoms of confusion on understanding JavaScript code and contribute to the design of new program transformation tools that aim to simplify JavaScript code.
\end{abstract}

@@

\input{1_Introduction}
\input{2_Background}
\input{3_Method}
\input{4_Result}
\input{discussion}
\input{5_Threats}
\input{6_Conclusion}
% \input{7_Atoms}

>>>>

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written or uses confuse idioms and language constructs, which might increase maintenance efforts and impose a significant cognitive load on developers. For this reason, researchers have investigated possible sources of confusion in codebases, including the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in statically typed languages such as \clang, \cpplang, and Java. In this work, we investigate whether atoms of confusion identified in statically typed languages also confuse developers of a dynamically typed language. We use JavaScript as a representative example of dynamic programming languages and collect evidence from a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. Our survey and interviews confirm that atoms of confusion lead to code that is hard to understand in JavaScript. Considering ten atom candidates, developers correctly predict the outcome of at least 15\% more cases in code snippets where atom candidates are not present, compared to alternative versions that include the atom candidates. For five of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for four atoms and medium for one of them. In addition, our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects, and two of them occur more than once for every 100 lines of code in the dataset. Altogether, our findings might help practitioners: (1) better understand the implications of atoms of confusion on understanding JavaScript code, (2) avoid writing code that is unnecessarily difficult to maintain, and (3) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
\end{abstract}

@@

\input{1_Introduction_alt}
\input{2_Background}
\input{3_Method}
\input{4_Result}
\input{discussion}
\input{5_Threats}
\input{6_Conclusion}
% \input{7_Atoms}
